---
title: "5_fieldsamples"
output: html_document
date: "2024-11-13"
---

load libraries
```{r, message=FALSE}
library(tidyverse)
library(dplyr)
rename <- dplyr::rename
```

load decontaminated reads and filter to retain just the field samples 
```{r}
decontaminated_table <- read.csv("/home/kimberly.ledger/nearshore_eDNA/outputs/samples_w_mocks_updated/decontaminated_asv_table.csv") %>%
  select(!X) %>%
  filter(sample_type == "sample") %>%
  filter(reads > 0)
```

load sample metadata
```{r}
metadata <- read.csv("/home/kimberly.ledger/nearshore_eDNA/data/samples_w_mocks_updated/nearshore_metadata.csv") %>%
  rename(Sample_ID = sample_ID) %>%
  filter(sample_type == "sample")
```

how many reads in field samples belong to ASVs/taxa that were included in the mock communities? 
```{r}
asvs_in_mocks <- read.csv("/home/kimberly.ledger/nearshore_eDNA/outputs/samples_w_mocks_updated/asvs_in_mocks.csv") %>%
  select(!X) 
```

add column to decontaminated table to designate if that ASVs was in the mock communities 
```{r}
decontaminated_table <- decontaminated_table %>%
  mutate(in_mock = ifelse(ASV %in% asvs_in_mocks$ASV, "yes", "no"))
```

summary table 
```{r}
decontaminated_table %>%
  group_by(in_mock) %>%
  summarize(total_reads = sum(reads))
```

so about 3/4th of reads come from taxa in the mock communities! that's pretty good. 

assign asv's to taxons 
```{r}
taxonomy <- read.csv("/home/kimberly.ledger/nearshore_eDNA/outputs/samples_w_mocks_updated/taxonomy_20241112_collapsed.csv") %>%
  select(!X) %>%
  select(qseqid, taxon) %>%
  unique() %>%
  rename(ASV = qseqid) 
```

join to decontaminated ASV table
```{r}
decon_table_w_tax <-  decontaminated_table %>%
  left_join(taxonomy, by = "ASV")
```

summary table 
```{r}
decon_table_w_tax %>%
  group_by(in_mock, taxon) %>%
  summarize(total_reads = sum(reads)) %>%
  pivot_wider(names_from = in_mock, values_from = total_reads)
```


based on this, i'll considier going back and make a few changes to the taxonomic assignments 
1 - put all Ammodytes reads to Ammodytes personatus (it's the only species present in region)
2 - lump the few Anoplarchus purpurescens reads to genus level 
3 - lump the few Cottus cognatus reads to genus level
4 - lump the few Myoxocephalus stelleri to genus level
5 - put Pholis reads to Pholis laeta

convert to taxon table 
```{r}
taxon_table <- decon_table_w_tax %>%
  select(Sample_ID, taxon, reads) %>%
  mutate(reads = as.integer(reads)) %>%
  pivot_wider(names_from = taxon, values_from = reads, values_fn = sum, values_fill = 0)
```

need to define project objectives before exploring data any more... 


